---
title: "Data Visulization DA2"
author: "Hima Rani Mathews 19BCE1532"
date: '2022-04-12'
output: html_document
---

## IMPORTING DATASET
```{r}
data<-read.csv("diabetes.csv")
head(data,10)
str(data)
dim(data)
summary(data)
```


### Finding if there is any null values
```{r}
any(is.na(data))
```

## DATA PREPROCESSING NAD CLEANING
### Count the number of 0 values in Glucose, Blood Pressure, Skin Thickness, BMI, and Insulin
```{r}
library(dplyr)
data %>%
filter(Glucose==0)%>%
count()
data %>% 
filter(BloodPressure==0) %>%
count()
data %>%
filter(SkinThickness==0)%>%
count()
data %>% 
filter(BMI==0) %>%
count()
data %>% 
filter(Insulin==0) %>%
count()
```


### Lots of cell have value as '0' hence replacing it with Median of each attribute
```{r}

diabetes<-data %>%
mutate(Glucose = ifelse(Glucose == 0, 117.0, Glucose), 
       BloodPressure= ifelse(BloodPressure==0, 72.00, BloodPressure),
       BMI= ifelse(BMI ==0,32.00, BMI),
       SkinThickness= ifelse(SkinThickness ==0, 23.00, SkinThickness),
       Insulin = ifelse(Insulin == 0, 30.5, Insulin))
```

### Changing the label names
```{r}
diabetes$DiabetesResult <- factor(diabetes$Outcome, levels = c("0", "1"), labels = c("negative", "positive"))
head(diabetes)
```

## VISULIZATIONS
```{r}
diabetes_result <- diabetes %>%
group_by (DiabetesResult) %>%
count(DiabetesResult)
diabetes_result
```

```{r}
library(ggplot2)
ggplot(data=diabetes_result) + geom_col(mapping=aes(x=DiabetesResult, y= n, fill=DiabetesResult)) +labs(title="Number of positive and negative diabetes cases", x= "Outcome", y = "Count")
```


## MULTIVARIATE ANALYSIS
### (a) PAIRPLOT
```{r}
pairs(diabetes[, 1:8],col=diabetes$DiabetesResult,main="PairPlot")

```

### (b) GGPAIR
Well, Pregnancies, Insulin, DBF and Age having skewed distribution.
```{r}
library(GGally)
ggpairs(diabetes[, 1:8], aes(color = diabetes$DiabetesResult),title="correlogram with ggpairs()")  +theme_bw()
```

### (c) CORRELATION MATRIX
```{r}
library("ggcorrplot")
corr <- round(cor(diabetes[, 1:8]), 1)
ggcorrplot(corr,
           hc.order = TRUE, type = "lower",
           color = c("#FC4E07", "white", "#00AFBB"),
           outline.col = "white", lab = TRUE)
```

### (d)SCATTERPLOT MATRIX
```{r}
library(lattice)
splom(~diabetes[, 1:8], col=diabetes$DiabetesResult, data = diabetes)
```

### (e)Principal component analysis
```{r}
library("factoextra")
my_data <- diabetes[, 1:8] 
res.pca <- prcomp(my_data, scale = TRUE)
fviz_pca_biplot(res.pca, col.ind = diabetes$DiabetesResult,
                palette = "jco", geom = "point")
```

### (f)Cluster analysis
```{r}
library(factoextra)
library(pheatmap)
diabetes[, 1:8]  %>%
   scale() %>%                  # Scale variables
  t() %>%                      # Transpose 
  pheatmap(cutree_cols = 4) 
```


